Grappa: Scalable Computing on Commodity Hardware \\for Irregular Applications

### 2012-07-23 doc/asplos13/evaluation.tex
tiny change
scaling v2

\subsection{Scaling}
% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. Grappa results are for 5 cores per node. Grappa with 20 machines is faster than the entire XMT of 128 processors.
%Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than Grappa. Fundamentally, aggregation as it is done in the current implementation of Grappa will not scale indefinitely, as increasing the number of destinations (individual cores across the cluster) does not scale because it requires more workers or time to form large enough messages.
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}

>>>>

\subsection{Scaling}
% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. Grappa results are for 5 cores per node. Grappa with 20 machines is faster than the entire XMT of 128 processors.
%Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than Grappa. We attribute this to a limitation in the current aggregator design and network stack that Grappa uses.  This limits the practical number of cores we can use to 6 per node (adding more cores per node \emph{decreases} performance).  Ironically, this limitation makes Grappa applications compute-bound instead of network-bound.  Work is ongoing to rework the infiniband driver stack and aggregation interface to it, to remove this limitation.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}
